\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-04-20 doc/paper.Rmd
results from cs2

errors2 <- rbind(errorrate(allp, 0.001), 
                errorrate(allp, 0.005), 
                errorrate(allp, 0.01), 
                errorrate(allp, 0.05))
errors2$type <- "profiles"

@@

Using the same notation as in \autoref{eq.start}, we define same-shift pairs identical to \citet{hadler} as pairs $(s_i^{(1)}, s_i^{(2)})$ for all $i$ where the boundary conditions of both sequences are met simultaneously. Let us assume that this results in $I$ pairs. Define $s_(j)^{(k)}$ to be the $j$th starting location in sequence $k = 1, 2$, i.e.\ $s_(1)^{(k)} < s_(2)^{(k)} < ... < s_(I)^{(k)}$.

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation}
i.e.\ for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
\end{figure}

@@

Profiles for each bullet land were extracted from  scans as described in \citet{aoas}.

## Setup
\hh{XXX the description below needs to be updated.}

\hh{using (a) signatures and (b) profiles, run chumbley score across scans from NIST and CSAFE for various settings of $w_o$ and $w_v$ (and coarseness $c$ for profiles).}

We used the adjusted Chumbley method as proposed in \citet{hadler} and implemented in the R package `toolmaRk` \citep{toolmark} on all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.

\hh{XXX also calculate results for the adjusted adjusted score XXX}
<!--The settings for optimizing and validating window sizes, $w_o$ and $w_v$, ranged from $w_o \in [50, 280]$ and $w_v \in \{30, 50\}$, see also figure \ref{fig:type2}.-->

## Failed Tests

Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$. \autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$. For same-source lands up to 12.5 percent of the tests fail.

```{r fails, echo=FALSE, fig.height = 3, fig.width = 5, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails. The percentage of failures is higher for same-source lands than for different source lands."}
fails <- allp %>% filter(coarse == 0.25) %>% 
  group_by(wo, wv) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n()
)

allp %>% filter(coarse == 0.25) %>% 
  group_by(wo, wv, match) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n()
) %>% 
  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
  geom_point(size=3) +
  xlab("Size of validation window in pixels") +
  ylab("Percent of failed tests") +
  theme_bw() +
  scale_colour_brewer("Same-source", palette="Set1") +
  scale_shape_discrete("Same-source") +
  ylim(c(0,13)) 
```

# A modified test

\hh{fix the algorithm - all of the other properties in Hadler and morris still hold.}

\hh{number of failed tests down. }

# Comparison of Results

## Profiles

\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}

\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness. The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minim in type II reached at a coarseness $c = 0.25$. }
\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
  geom_point() + geom_line() +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
  ylab("Type II error") 
e1
```

@@

```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top). The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$. "}
profile <- sigs_graphics[[1]] %>% filter(land_id==19)
profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
profile$lw30 <- lowess(x = profile$y, y = profile$value, f=.30)$y
#profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
#profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
#profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y

profiles <- tidyr::gather(profile, coarseness, values, lw20:lw30)
profiles <- profiles %>% mutate(
  c = readr::parse_number(coarseness)/100
)
p1 <- profiles %>% 
  ggplot(aes(x = y, y = value-values)) + geom_line() +
  theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
  facet_grid(~c, labeller="label_both")

@@

## Results

>>>>

errors2 <- rbind(errorrate(allp, 0.001), 
                errorrate(allp, 0.005), 
                errorrate(allp, 0.01), 
                errorrate(allp, 0.05))
errors2$type <- "CS1"

if (!file.exists("../data/invprofiles.rds")) {
  files <- dir("../data/inverse-profiles/", pattern="rds")
  allinv <- data.frame()
  for (file in files) {
    tmp <- readRDS(file=file.path(path="../data/inverse-profiles", file)) 
    allinv <- rbind(allinv, tmp)
  }
  allinv <- allinv %>% filter(land1_id < land2_id)
  allinv$coarse <- allinv$coarseness
  allinv <- allinv %>% select(-coarseness)
  saveRDS(allinv, file="../data/invprofiles.rds")
} else {
  allinv <- readRDS("../data/invprofiles.rds")
}

errors3 <- rbind(errorrate(allinv, 0.001), 
                errorrate(allinv, 0.005), 
                errorrate(allinv, 0.01), 
                errorrate(allinv, 0.05))
errors3$type <- "CS2"

@@

Using the same notation as in \autoref{eq.start}, we define same-shift pairs identical to \citet{hadler} as pairs $(s_i^{(1)}, s_i^{(2)})$ for all $i$ where the boundary conditions of both sequences are met simultaneously. Let us assume that this results in $I$ pairs. Define $s_{(j)}^{(k)}$ to be the $j$th starting location in sequence $k = 1, 2$, i.e.\ $s_{(1)}^{(k)} < s_{(2)}^{(k)} < ... < s_{(I)}^{(k)}$.

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation}
i.e.\ for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.
\hh{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as {\bf (CS1)} and the suggested modified algorithm as {\bf (CS2)} and compare their performance on the available scans of the Hamby study.}
\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
\end{figure}

@@

Profiles for each bullet land were extracted from scans close to the heel of the bullet while avoiding break-off as described in \citet{aoas}.

## Setup
Both algorithms (CS1) and (CS2) are implemented in the R package `toolmaRk` \citep{toolmark}. We applied both methods to all pairwise land-to-land comparisons of the Hamby scans provided by NIST for a total of 85,491 land-to-land comparisons.

# Results
## Failed Tests

Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$. 
\autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$. For same-source lands up to 12.5 percent of the tests fail using CS1. The highest percentage of failed tests under CS2 is 0.9% for different source tests using a validation window size $w_v$ of 50 pixels. }

```{r fails, echo=FALSE, fig.height = 3, fig.width = 7, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2)."}
allinv$method <- "CS2"
allp$method <- "CS1"

prof_fails <- rbind(
  allinv %>% select(wo, wv, method, p_value, match),
  allp %>% filter(coarse == 0.25) %>% 
    select(wo, wv, method, p_value, match)
)
fails <- prof_fails %>%
  group_by(wo, wv, method, match) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n()
)

fails %>% 
  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
  geom_point(size=3) +
  xlab("Size of validation window in pixels") +
  ylab("Percent of failed tests") +
  theme_bw() +
  scale_colour_brewer("Same-source", palette="Set1") +
  scale_shape_discrete("Same-source") +
  ylim(c(0,13)) +
  facet_wrap(~method, labeller="label_method")
```

\hh{Coarseness is another parameter that has to be considered in the matching of bullet land profiles.}

@@

```{r profile-sketch, fig.height = 5, fig.width='\\textwidth', fig.cap="Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \\ref{fig:sigs-profiles} (top). The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$. "}
profile <- sigs_graphics[[1]] %>% filter(land_id==19)
profile$lw10 <- lowess(x = profile$y, y = profile$value, f=.10)$y
profile$lw15 <- lowess(x = profile$y, y = profile$value, f=.15)$y
profile$lw20 <- lowess(x = profile$y, y = profile$value, f=.20)$y
profile$lw25 <- lowess(x = profile$y, y = profile$value, f=.25)$y
profile$lw30 <- lowess(x = profile$y, y = profile$value, f=.30)$y
#profile$lw50 <- lowess(x = profile$y, y = profile$value, f=.5)$y
#profile$lw75 <- lowess(x = profile$y, y = profile$value, f=.75)$y
#profile$lw100 <- lowess(x = profile$y, y = profile$value, f=1)$y

profiles <- tidyr::gather(profile, coarseness, values, lw10:lw30)
profiles <- profiles %>% mutate(
  c = readr::parse_number(coarseness)/100
)
p1 <- profiles %>% 
  ggplot(aes(x = y, y = value-values)) + geom_line() +
  theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) +
  facet_grid(~c, labeller="label_both")

@@

\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness. The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$. }
\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
  geom_point() + geom_line() +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
  ylab("Type II error") 
e1
```

### Results for Signatures

### 2018-04-20 doc/paper.tex
setup for both methods
results from cs2

Using the same notation as in \autoref{eq.start}, we define same-shift
pairs identical to \citet{hadler} as pairs \((s_i^{(1)}, s_i^{(2)})\)
for all \(i\) where the boundary conditions of both sequences are met
simultaneously. Let us assume that this results in \(I\) pairs. Define
\(s_(j)^{(k)}\) to be the \(j\)th starting location in sequence
\(k = 1, 2\), i.e.~\(s_(1)^{(k)} < s_(2)^{(k)} < ... < s_(I)^{(k)}\).

@@

i.e.~for an odd number of same-shift correlations, we skip the middle
pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the
same or at most one less than the number of same-shift pairings in all
tests.

@@

\hh{Discuss CSAFE scans}

Profiles and signatures were extracted from all scans as described in
\citet{aoas}.

@@

\hh{XXX the description below needs to be updated.}

\hh{using (a) signatures and (b) profiles, run chumbley score across scans from NIST and CSAFE for various settings of $w_o$ and $w_v$ (and coarseness $c$ for profiles).}

We used the adjusted Chumbley method as proposed in \citet{hadler} and
implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
pairwise land-to-land comparisons of the Hamby scans provided by NIST
for a total of 85,491 land-to-land comparisons.

@@

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail.

@@

\section{A modified test}\label{a-modified-test}

\hh{fix the algorithm - all of the other properties in Hadler and morris still hold.}

\hh{number of failed tests down. }

\section{Comparison of Results}\label{comparison-of-results}

\subsection{Profiles}\label{profiles}

\hh{When dealing with profiles, coarseness is an additional parameter that has to be considered in the matching.}

\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness. The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minim in type II reached at a coarseness $c = 0.25$. }
\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}

@@

\subsection{Results}\label{results}

>>>>

Using the same notation as in \autoref{eq.start}, we define same-shift
pairs identical to \citet{hadler} as pairs \((s_i^{(1)}, s_i^{(2)})\)
for all \(i\) where the boundary conditions of both sequences are met
simultaneously. Let us assume that this results in \(I\) pairs. Define
\(s_{(j)}^{(k)}\) to be the \(j\)th starting location in sequence
\(k = 1, 2\),
i.e.~\(s_{(1)}^{(k)} < s_{(2)}^{(k)} < ... < s_{(I)}^{(k)}\).

@@

i.e.~for an odd number of same-shift correlations, we skip the middle
pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the
same or at most one less than the number of same-shift pairings in all
tests.
\hh{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as {\bf (CS1)} and the suggested modified algorithm as {\bf (CS2)} and compare their performance on the available scans of the Hamby study.}

@@

Profiles for each bullet land were extracted from scans close to the
heel of the bullet while avoiding break-off as described in
\citet{aoas}.

@@

Both algorithms (CS1) and (CS2) are implemented in the R package
\texttt{toolmaRk} \citep{toolmark}. We applied both methods to all
pairwise land-to-land comparisons of the Hamby scans provided by NIST
for a total of 85,491 land-to-land comparisons.

\section{Results}\label{results}

@@

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail using CS1.
\hh{XXX Currently running CS2 for these settings for a comparison.}

@@

\hh{XXX Conclusion: the number of failed tests goes down - a lot. XXX}

\hh{Coarseness is another parameter that has to be considered in the matching of bullet land profiles.}

@@

\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness. The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$. }
\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}

\subsubsection{Results for Signatures}\label{results-for-signatures}
